---
title: "Exploring Healthcare-Associated Infections (HAI) Burden with BHAIpkgHafsa"
author: "Hafsa Altaf"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{BHAIpkgHafsa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

## Overview

The BHAIpkgHafsa package provides an interactive tool for exploring the burden of Healthcare-Associated Infections (HAIs) across different infection types.

It includes:

- A clean, ready-to-use dataset (healthcare_associated_infections)

- A fully interactive Shiny app for visualizing and comparing HAI burden

- Supporting documentation and reproducible examples

This vignette demonstrates how to load the dataset, inspect it, and use the interactive app.

### 1. Installation

You can install the package directly from your GitHub repository using the remotes package:


```{r setup}
# Install remotes if not already installed
# install.packages("remotes")

# Install BHAIpkgHafsa from GitHub (replace with your repo URL)
# remotes::install_github("yourusername/BHAIpkgHafsa")

# Load the package
# library(BHAIpkgHafsa)
```

### 2. Dataset: healthcare_associated_infections

The dataset included in this package contains estimates of healthcare-associated infections (HAIs) by infection type, including both case counts and attributable deaths.

#### Load dataset

data("healthcare_associated_infections", package = "BHAIpkgHafsa")

#### View structure

str(healthcare_associated_infections)

##### Example rows

head(healthcare_associated_infections)

##### Variables included/Variable	Description

- Infection_Type:	The category of healthcare-associated infection (e.g., BSI, HAP, UTI, SSI, CDI)
- Number_of_HAIs:	Estimated total number of infection cases occurring in healthcare settings
- Number_of_attributable_deaths: Estimated number of deaths directly caused by each infection type

### 3. Launching the Interactive Shiny App

The heart of the package is the interactive Shiny dashboard, which allows users to:

- Select an infection type (or compare all types)

- Choose the metric to visualize (HAIs vs. attributable deaths)

- View interactive bar plots with hover tooltips

- Display a dynamic summary table

Launch the app using a single command:

```{r}
# launch_haiexplorer()
```

This will open a browser window displaying the HAI Burden Explorer dashboard.

#### Features of the App

- Interactive Plotly Visualization:
Hover over the bars to see details of infection burden by type.

- Dynamic Filters:
Use the dropdown and radio buttons to explore specific infections or compare across all.

- Summary Table:
Toggle to view numeric summaries of HAI and death counts.

- Styled Interface:
The app uses a modern blue theme, rounded panels, and soft background for readability.

### 4. Example Exploration Workflow

Here’s a typical exploration sequence:

1. Launch the app:

```{r}
# launch_haiexplorer()
```

2. In the dashboard:

- Select “All” to compare all infection types.

- Switch between Number of HAIs and Number of attributable deaths.

- Hover over bars to read tooltip information.

- Enable the “Show summary table” checkbox to view a numeric breakdown.

3. Interpretation:

- Infection types with taller bars indicate higher burden.

- Comparing HAI counts and attributable deaths reveals severity differences among infections.

### 5. Package Structure

Below is the internal structure of the BHAIpkgHafsa package:


```{r}
# BHAIpkgHafsa/
# ├── DESCRIPTION
# ├── NAMESPACE
# ├── R/
# │   └── launch.R
# ├── inst/
# │   └── shiny/
# │       └── app.R
# ├── data/
# │   └── healthcare_associated_infections.rda
# ├── man/
# │   └── launch_haiexplorer.Rd
# └── vignettes/
#     └── Using_BHAIpkgHafsa.Rmd
```

### 6. Acknowledgements

This package and dashboard were developed as part of an assessment project to demonstrate the creation of an R package with an integrated Shiny app and pkgdown site.

Special thanks to the R community and the authors of tidyverse, plotly, shiny, and usethis for enabling modern, reproducible data visualization.



